Meta-analysis of composite outcomes in clinical trials

### 2025-01-28 MSc_poster_template.Rmd
Background and Problem updated

# Background and Problem

This template has been produced using the `posterdown` package. It generates a standard conference-style poster layout.

<!-- You can add references by inserting the following ^[reference details here] -->

A common problem in clinical research is messed up data[^1]. It can lead to bias or nonsensical results. Appropriate statistical analysis can result in superb findings in such cases, with the right assumptions.

[^1]: Evans 2010, doi: 10.6030/1939-067x-3.1.1

## Objectives of Project

1.  Explain the statistical issue or research question in health data science to be investigated.
2.  Explore a method of analysis.
3.  Apply any methods to selected datasets.

@@

You can even use the bibliography the same way: Our data were taken from a cluster-randomised trial, available from the [Irish Social Science Data Archive](https://www.ucd.ie/issda/data/).

@@

# Early Results 
```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite. Test statistic 28.89, P-value of P<0.001', fig.height=4.5}
continuous_permutation_function<-function(x){
  permutation_t_value<-numeric(10000)
  i <-1
  repeat{
    thesis.df$Group<-sample(thesis.df$Group, 535, replace = F)
    permutation_result<-t.test(x[thesis.df$Group == "Metformin"], x[thesis.df$Group == "Placebo"])
    permutation_t_value[i]<-permutation_result$statistic
    i<- i +1
    if(i > 10000) break}
  return(permutation_t_value)
}
squared_composite<-((continuous_permutation_function(thesis.df$binary_insulin))^2) +
  ((continuous_permutation_function(thesis.df$gw32))^2) +
  ((continuous_permutation_function(thesis.df$gw38))^2)
percentile_95<-quantile(squared_composite, 0.95)

>>>>

# Background and Problem
Composites are commonly used in clinical trials as the primary outcome. When misused, they can reduce the power to detect treatment effect. This was seen in EMERGE, a randomised clinical trial investigating the effects of early initiation of Metformin treatment on developing gestational diabetes. The composite was created using three measures of glycemic control; blood glucose levels of 5.1mmol/L or greater at gestation weeks 32 or 38, or initiation of Insulin treatment.

 ***Insulin~Yes~ OR gw32≥5.1~Yes~ OR gw38≥5.1~Yes~ ***

This composite involves dichotomising blood glucose levels, greatly reducing power. Therefore, the objectives of this project are:

1.	To perform a meta-analysis of the primary outcome using permutations of the composite test statistic, and analyzing the glucose components as continuous variables

2.	To compare various ways of creating the composite test statistic, and their power

3.	To compare the permutation method with other approaches to analyzing composite outcomes

@@

@@

# Early Results 
```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite. Test statistic 28.89, *p*-value of *p*<0.001', fig.height=4.5}
continuous_permutation_function<-function(x){
  permutation_t_value<-numeric(10000)
  i <-1
  repeat{
    thesis.df$Group<-sample(thesis.df$Group, 535, replace = F)
    permutation_result<-t.test(x[thesis.df$Group == "Metformin"], x[thesis.df$Group == "Placebo"])
    permutation_t_value[i]<-permutation_result$statistic
    i<- i +1
    if(i > 10000) break}
  return(permutation_t_value)
}
squared_composite<-((continuous_permutation_function(thesis.df$binary_insulin))^2) +
  ((continuous_permutation_function(thesis.df$gw32))^2) +
  ((continuous_permutation_function(thesis.df$gw38))^2)
percentile_95<-quantile(squared_composite, 0.95)
